---
title: "02-standardize_plates"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# packages

```{r,echo=TRUE,warning=FALSE,message=FALSE}
library(tidyverse)
```

# import plates in long-format

```{r}
data4 <- read_rds("data/multiplate-long.rds")

data4 %>% 
  # epihelper::print_inf() %>% 
  identity()
```

# create standardization dataframe

```{r}
std_reference <- 
  tribble(
    ~dilution,~reciprocal,~factor,~abunits,
    "STD 1/50",50,"102400/50",NA_real_,
    "STD 1/100",100,"102400/100",NA_real_,
    "STD 1/200",200,"102400/200",NA_real_,
    "STD 1/400",400,"102400/400",NA_real_,
    "STD 1/800",800,"102400/800",NA_real_,
    "STD 1/1600",1600,"102400/1600",NA_real_,
    "STD 1/3200",3200,"102400/3200",NA_real_,
    "STD 1/6400",6400,"102400/6400",NA_real_,
    "STD 1/12800",12800,"102400/12800",NA_real_,
    "STD 1/25600",25600,"102400/25600",NA_real_,
    "STD 1/51200",51200,"102400/51200",NA_real_,
    "STD 1/102400",102400,"102400/102400",NA_real_,
  ) %>% 
  mutate(abunits=max(reciprocal)/reciprocal)

std_reference
```

# build density ~ abunit relationship

## isolate standard curves

```{r}
data5 <- data4 %>% 
  filter(str_detect(codigo,"STD")) %>% 
  left_join(std_reference,by = c("codigo"="dilution"))

data5 %>% 
  # epihelper::print_inf() %>%
  identity()
```

```{r}
data6 <- data4 %>% 
  filter(str_detect(codigo,"ctrl"))
```

## descriptive

```{r,fig.height=10,fig.width=10}
data5 %>% 
  ggplot(aes(x = abunits,y = densidad)) +
  geom_point() +
  # geom_smooth() +
  geom_hline(aes(yintercept=densidad, color = codigo),
             data = data6) +
  scale_x_log10() +
  facet_wrap(~plate+experimento)
```

## log-logistic model

- [pending]

```{r}

```

# quality control of replicates

## replicates per plate

- how many sample replicates are in each plate?

```{r,message=FALSE}
data4 %>% 
  filter(!str_detect(codigo,"STD|ctrl")) %>% 
  count(plate,experimento,codigo,sort = T) %>% 
  count(n)

data4 %>% 
  filter(!str_detect(codigo,"STD|ctrl")) %>% 
  count(plate,experimento,codigo,sort = T) %>% 
  # epihelper::print_inf() %>% 
  identity()
```

## replicates all plate

- how many sample replicates are in all plates?

```{r,message=FALSE}
data4 %>% 
  filter(!str_detect(codigo,"STD|ctrl")) %>% 
  count(codigo,sort = T) %>% 
  count(n)

data4 %>% 
  filter(!str_detect(codigo,"STD|ctrl")) %>% 
  count(codigo,sort = T)
```

## replicate variability

```{r}
data7 <- data4 %>% 
  filter(!str_detect(codigo,"STD|ctrl")) %>% 
  group_by(plate,experimento,codigo) %>% 
  summarise(densidad_mean=mean(densidad),
            densidad_sd=sd(densidad)) %>% 
  ungroup() %>% 
  mutate(densidad_cv=100*densidad_sd/densidad_mean) %>% 
  # epihelper::print_inf() %>% 
  identity()

data7
```

### identify outliers

```{r}
data8 <- data7 %>% 
  filter(densidad_mean>900)

data8
```

```{r}
data8 %>% 
  select(plate,codigo)
data4 %>% 
  filter(magrittr::is_in(
    x = plate,
    table = data8 %>% pull(plate))) %>% 
  filter(magrittr::is_in(
    x = codigo,
    table = data8 %>% pull(codigo)))
```


### distribution table

#### con outlier

```{r}
data7 %>% 
  select(starts_with("densidad")) %>% 
  skimr::skim() %>% 
  skimr::yank(skim_type = "numeric")
```

#### sin outlier

```{r}
data7 %>% 
  filter(densidad_mean<900) %>% 
  select(starts_with("densidad")) %>% 
  skimr::skim() %>% 
  skimr::yank(skim_type = "numeric")
```

### distribution plot

#### con outlier

```{r,fig.height=10,fig.width=10}
data7 %>% 
  ggplot(aes(x = densidad_mean,y = densidad_cv)) +
  geom_point() +
  geom_hline(aes(yintercept=20),
             linetype="dashed",size=0.3) +
  geom_vline(aes(xintercept=0.25),
             linetype="dashed",size=0.3) +
  coord_cartesian(ylim = c(0,100)) +
  facet_wrap(~plate+experimento)
```

#### sin outlier

```{r,fig.height=10,fig.width=10}
data7 %>% 
  filter(densidad_mean<900) %>% 
  ggplot(aes(x = densidad_mean,y = densidad_cv)) +
  geom_point() +
  geom_hline(aes(yintercept=20),
             linetype="dashed",size=0.3) +
  geom_vline(aes(xintercept=0.25),
             linetype="dashed",size=0.3) +
  coord_cartesian(ylim = c(0,100)) +
  facet_wrap(~plate+experimento)
```

# estimate abunits of samples

- [pending]

```{r}

```

# view plate shape

```{r,eval=FALSE}
plater::view_plate(
  data = data4 %>% filter(plate=="placa_001"), 
  well_ids_column = "pozo", 
  columns_to_display = c("codigo","densidad")
)

plater::view_plate(
  data = data4 %>% filter(plate=="placa_002"), 
  well_ids_column = "pozo", 
  columns_to_display = c("codigo","densidad")
)
```


# end

```{r}

```

# draft

### 4pll per template

```{r,message=FALSE, eval=FALSE}

mab_ir <- NULL
mod_bx <- NULL

#
# new dose levels as support for the line
#mdo$Ab.units %>% summary()
new_x <- expand.grid(exp(seq(log(0.1),log(2048),length=100)))
# db to add predictions of all plates
new <- data_frame(ord=as.character(),
                  resp=as.double(),
                  p=as.double(),
                  pmin=as.double(),
                  pmax=as.double(),
                  Plate=as.character())
#

for (j in 1:length(levels(phe$Plate))) {
  
#
# 5 PARAMETER ESTIMATION 4pLL model
#
wb.m1 <- drm(OD ~ Ab.unit, Plate, 
               data= std %>% filter(Plate==levels(phe$Plate)[j]),#j
             #data= std,
               fct = LL.4(names = c("b", "c", "d", "e")))
#
wb.model <- wb.m1
# 6 BOX-COX TRANSFORMATION against RESIDUAL heterogeneity
wb.model.BX <- boxcox(wb.model, 
                     main=expression("Optimal " ~ lambda ~ " with confidence intervals"), 
                     plotit = FALSE)
#coefficients(wb.model.BX) %>% matrix(7,4)
mab_p <- mab %>% as.data.frame()
# 7 UNK AB.UNITS ESTIMATION by INVERSE REGRESSION
mir <- ED(wb.model.BX, 
                 mab_p[mab_p$Plate==levels(phe$Plate)[j],"mean.OD"],#j
                   #wb_MEAN[1:n,5],
                   type = "absolute",interval = "delta",
                   #clevel = "Pfal", 
                 display = FALSE)
  
mab_ir <- rbind(mab_ir,mir)
mod_bx <- rbind(mod_bx,coefficients(wb.model.BX))

#
# predictions and confidence intervals
pdm <- predict(wb.model.BX, newdata = new_x, interval = "confidence")
# new data with predictions
new_p <- bind_cols(new_x %>% 
                     as.tibble() %>% 
                     rownames_to_column(var = "ord")
                   , pdm %>% 
                     as.tibble() %>% 
                     rownames_to_column(var = "ord")
                   ) %>%
  dplyr::select(-ord1) %>% 
  mutate(Plate=levels(phe$Plate)[j]) %>% 
  dplyr::rename(resp=Var1,p=Prediction,pmin=Lower,pmax=Upper)

new <- union(new,new_p)
#

}

#mab
#mab[mab$Plate==levels(phe$Plate)[1],] #%>% duplicated() %>% sum()

# 7.1 FEED UNK AB.UNITS DATA.FRAME
mdo <- mab_ir %>% as.data.frame() %>% rownames_to_column() %>% 
  #dplyr::rename(ord=rowname) %>% 
  separate(rowname,c("par","Plate","mean.OD.c"),sep = ":") %>% 
  rownames_to_column("ord") %>% 
  #mutate(ord=seq(1,dim(.)[1])) %>% 
  full_join(mab %>% 
              mutate(ord=as.character(ord)) %>% 
              mutate(mean.OD.c=as.character(mean.OD))
            ,
            by = c("ord","Plate","mean.OD.c")) %>% 
  dplyr::select(Plate,ord,ID,code,Type,#pheno,igg,
                mean.OD,sd.OD,cv.OD,Ab.units=Estimate,
         everything(),-par,-mean.OD.c,-Ab.unit) #%>% 
  # ANY MANUAL FILTERING of replicate on different templates could be applied HERE!!!
  #filter(!code=="2235_137") 

mod_bt <- mod_bx %>% as.data.frame() %>% rownames_to_column() %>% as.tibble() %>% 
  full_join(std %>% 
              dplyr::count(Plate) %>% dplyr::select(-n) %>% rownames_to_column(),
            by="rowname") %>% 
  dplyr::select(Plate,everything(),-rowname) %>% 
  separate(Plate,c("Plate","anti"),sep = "_") #%>% 
  #dplyr::rename(Plate=rowname) %>% 
  #mutate(Plate=stringr::str_replace(Plate,"(\\d)","N\\1"))

new <- new %>% mutate(ord=as.numeric(ord)) %>% arrange(Plate,ord)
```

### outputs

```{r, eval=FALSE}
#
#fin
#end
#mod
#mab

# standard curve data
std

# estimated ab unit data
mdo

# estimated parameters per standard curve
mod_bt

# predicted model per standard curve 
new
```
