---
title: "03-estimate_abunits"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# packages

```{r,echo=TRUE,warning=FALSE,message=FALSE}
if(!require(pak)) install.packages("pak")
pak::pak(tidyverse)
pak::pak(drc)
```

```{r}
options(scipen=10000)
```


# import input data

```{r}
data_stdc <- read_rds("data/02-data-standard_curve.rds")
data_ctrl <- read_rds("data/02-data-only_control.rds")
data_abmean <- read_rds("data/02-data-ab_mean.rds") 
```

# fit 4pll model

```{r}
# mab_ir <- NULL
# mod_bx <- NULL
j <- 1
```


## get data

```{r}
data_drm_stdc <- data_stdc %>% 
  filter(plate==unique(data_stdc$plate)[j]) %>% 
  filter(experimento==unique(data_stdc$experimento)[j]) %>% 
  # mutate(plate=as.factor(plate)) %>% 
  # as.data.frame() %>% 
  identity()
# data_drm_stdc %>% class()
# data_drm_stdc
```

```{r}
data_drm_ctrl <- data_ctrl %>% 
  filter(plate==unique(data_ctrl$plate)[j]) %>% 
  filter(experimento==unique(data_ctrl$experimento)[j]) %>% 
  # mutate(plate=as.factor(plate)) %>% 
  # as.data.frame() %>% 
  identity()
```

```{r}
data_drc_abmean <- data_abmean %>% 
  filter(plate==unique(data_ctrl$plate)[j]) %>% 
  filter(experimento==unique(data_ctrl$experimento)[j]) %>% 
  # mutate(plate=as.factor(plate)) %>% 
  # as.data.frame() %>% 
  identity()
```


## dose response model

```{r}
# one plate experiment
wb.model <- 
  drm(formula = densidad ~ abunits, 
    curveid = plate,
    data= data_drm_stdc,
    fct = LL.4(names = c("b", "c", "d", "e")))
#' "b" -> "Slope", 
#' "c" -> "Lower Limit", 
#' "d" -> "Upper Limit", 
#' "e" -> "ED50"
```

## tidy model output

```{r}
wb.model %>% broom::tidy()
wb.model %>% broom::glance()
# wb.model %>% pluck("data")
```

## box-cox transformation

BOX-COX TRANSFORMATION against RESIDUAL heterogeneity

```{r,warning=FALSE}
# only one plate

wb.model.BX <- 
  boxcox(object = wb.model, 
         # main=expression("Optimal " ~ lambda ~ " with confidence intervals"), 
         plotit = FALSE)
```

## tidy model output

```{r}
wb.model.BX %>% broom::tidy()
```


If needed, choose the first model with:

```{r}
# wb.model.BX <- wb.model
```

# predict the standard curve

## prediction and confidence interval

```{r}
data_new <- expand.grid(exp(seq(log(0.1),log(2048),length=100)))
```

```{r,warning=FALSE}
# only one plate

# predictions and confidence intervals
pdm <- predict(object = wb.model.BX, newdata = data_new, interval = "confidence")

## join data
data_new_pred <- data_new %>% 
  rownames_to_column() %>% 
  left_join(
    pdm %>% 
      as_tibble() %>% 
      rownames_to_column()) %>% 
  mutate(plate=unique(data_stdc$plate)[j]) %>% 
  dplyr::rename(resp=Var1,p=Prediction,pmin=Lower,pmax=Upper) %>% 
  dplyr::select(plate,everything())

```

## result

```{r,fig.height=4,fig.width=6,fig.align='center'}
ggplot() +
  geom_hline(data = data_drm_ctrl,
             mapping = aes(yintercept=densidad, color = codigo)) +
    geom_ribbon(data = data_new_pred,
              mapping = aes(ymin = pmin, ymax = pmax, 
                            x = resp), fill = "grey70") +
  geom_line(data = data_new_pred,
            mapping = aes(x = resp,y = p)) +
  geom_point(data = data_drm_stdc,
             mapping = aes(x = abunits,y = densidad)) +
  scale_x_log10() +
  facet_wrap(~plate+experimento) +
  labs(x = "Unidades de Anticuerpos",
       y = "Densidad óptica")
```

# estimate abunits

UNK AB.UNITS ESTIMATION by INVERSE REGRESSION [pending]

## estimates and confidence intervals

```{r,warning=FALSE}
# only one plate

data_ed_abunit <-
  ED(object = wb.model.BX,
     respLev = data_drc_abmean %>% pull(densidad_mean),
     #j #wb_MEAN[1:n,5],
     type = "absolute",
     interval = "delta",
     #clevel = "Pfal",
     display = FALSE)

data_abmean_abunit <-
  data_drc_abmean %>%
  rownames_to_column() %>%
  left_join(
    data_ed_abunit %>%
      as_tibble() %>%
      rownames_to_column() %>%
      janitor::clean_names()
  )
```

## result

```{r,fig.height=4,fig.width=6,fig.align='center'}
ggplot() +
  geom_hline(data = data_drm_ctrl,
             mapping = aes(yintercept=densidad, color = codigo)) +
  geom_ribbon(data = data_new_pred,
              mapping = aes(ymin = pmin, ymax = pmax, 
                            x = resp), fill = "grey70") +
  geom_line(data = data_new_pred,
            mapping = aes(x = resp,y = p)) +
  geom_point(mapping = aes(x = estimate,y = densidad_mean),
             data = data_abmean_abunit) +
  geom_errorbar(mapping = aes(x = estimate,y = densidad_mean,
                    xmin = lower, xmax = upper),
                data = data_abmean_abunit) +
  scale_x_log10() +
  facet_wrap(~plate+experimento) +
  labs(x = "Unidades de Anticuerpos",
       y = "Densidad óptica")
```

# outputs

```{r}
# data_abmean_abunit
```

# references

- review https://rstats4ag.org/dose-response-curves.html
- review https://bioassay.dk/