---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# elixr-casestudy01

<!-- badges: start -->
<!-- badges: end -->

The goal of `elixr-casestudy01` is to host the code used for a serological survey workflow to:

- __Collect__ multiple plate files in a single file using `{plater}`,
- __Standardize__ Antibody Units across multiple plates using method proposed by [Miura K, 2008](https://www.sciencedirect.com/science/article/pii/S0264410X07012601), and
- __Classify__ the seropositivity of the Antibody Units of the sample using method proposed by [Sepulveda N, 2015](https://www.hindawi.com/journals/jir/2015/738030/)


## Installation

* Clone this repository. Follow [this steps to create a new Rstudio Project from a GitHub repository](https://www.epirhandbook.com/en/version-control-and-collaboration-with-git-and-github.html?q=clone#clone-from-a-github-repository). 

* Restore the dependencies of this R project, with specific packages and versions, with `renv::restore()`.

* Run:

```{r eval = FALSE}
source("make.R")
```


## Contributing

See our [Contributing guide](/.github/CONTRIBUTING.md) for ways to get started.

Please adhere to this project's [Code of Conduct](/.github/CODE_OF_CONDUCT.md).


## Support

Please see our [Getting help guide](/.github/SUPPORT.md) for support.


## License

[GPLv3](https://choosealicense.com/licenses/gpl-3.0/)


## Authors

- [@avallecam](https://www.github.com/avallecam)

## Acknowledgements

Many thanks for the opportunity to work on this project to the team at:

Laboratorio de Biología Molecular e Inmunología, 
Unidad Especializada de Investigación en Productos Naturales Antiparasitarios de la Amazonia, 
Centro de Investigación en Recursos Naturales ([CIRNA](https://www.gob.pe/37761-universidad-nacional-de-la-amazonia-peruana-centro-de-investigaciones-de-recursos-naturales-de-la-unap-cirna-unap)), 
Universidad Nacional de la Amazonia Peruana ([UNAP](https://enlinea.unapiquitos.edu.pe/)),
Iquitos - Perú.

## References

Miura K, Orcutt AC, Muratova OV, Miller LH, Saul A, Long CA. Development and characterization of a standardized ELISA including a reference serum on each plate to detect antibodies induced by experimental malaria vaccines. Vaccine. 2008;26(2):193-200. doi:[10.1016/j.vaccine.2007.10.064](https://www.sciencedirect.com/science/article/pii/S0264410X07012601).

Sepúlveda N, Stresman G, White MT, Drakeley CJ. Current mathematical models for analyzing anti-malarial antibody data with an eye to malaria elimination and eradication. Journal of Immunology Research. 2015;2015:1-21. doi:[10.1155/2015/738030](https://www.hindawi.com/journals/jir/2015/738030/).

Saavedra-Langer, R., Marapara, J., Valle-Campos, A., Durand, S., Vásquez-Chasnamote, M. E., Silva, H., & Pinedo-Cancino, V. (2018). IgG subclass responses to excreted-secreted antigens of Plasmodium falciparum in a low-transmission malaria area of the Peruvian Amazon. Malaria journal, 17(1), 1-10. doi: [10.1186/s12936-018-2471-6](https://malariajournal.biomedcentral.com/articles/10.1186/s12936-018-2471-6).
